### 2020-04-21 thesis_template/3_Body/chapter/concept.tex
documented different variable and drop reference

Our KNN algorithms are implemented with batch processing. In query phase, we control batch size to examine how size of objects allocated simultaneously in memory.
The detail implementation is explained in section3.2.
\clearpage

>>>>

Our KNN algorithms are implemented with batch processing. In query phase, we control batch size to examine how size of objects allocated simultaneously in memory.
The detail implementation is explained in section3.2.

\subsection{Complex Objects}
\label{sec:history}
To conduct experiments for above concepts, we use the 4 types of complex object: CustomerOwned, CustomerBorrowed, CustomerSlice, and CustomerRc. 
These object contains other type of object: OrderOwned, OrderBorrowed, OrderSlice, and OrderRc.
The representation of these objects are shown in Figure~\ref{fig:customer} and Figure~\ref{fig:order}. 
Three Customer objects have 15 fields: 3 fields for i32, 3 fields for f64, 8 fields for String, and 1 field for Order object.
All of fields of CustomerOwned and OrderOwned are owned by the object. On the other hand, fields of CustomerBorrowed, CustomerSlice, OrderBorrowed, and OrderSlice are borrowed. 
Reference of slice of values are used as the fields and owner of actual values are stored differently in source Vec. 
CustomerRc acquires Rc of values used for its fields from the source Vec.

\begin{figure}[htb]
    \begin{minipage}[t]{0.2\linewidth}\centering
        \begin{lstlisting}
            struct CustomerOwned {
                key: i32,
                age: i32,
                num_purchase: i32,
                total_purchase: f64,
                duration_spent: f64, 
                duration_since: f64,
                zip_code: String,
                address: String,
                country: String,
                state: String,
                first_name: String,
                last_name: String,
                province: String,
                comment: String, 
                order: OrderOwned
            }
        \end{lstlisting}
      \medskip
      \centerline{(a)}
    \end{minipage}\hfill
    \begin{minipage}[t]{0.6\linewidth}\centering
        \begin{lstlisting}
            struct CustomerBorrowed<'a> {
                key: &'a i32,
                age: &'a i32,
                num_purchase: &'a i32,
                total_purchase: &'a f64,
                duration_spent: &'a f64, 
                duration_since: &'a f64,
                zip_code: &'a String,
                address: &'a String,
                country: &'a String,
                state: &'a String,
                first_name: &'a String,
                last_name: &'a String,
                province: &'a String,
                comment: &'a String, 
                order: &'a OrderBorrowed<'a>
            }
        \end{lstlisting}
      \medskip
      \centerline{(b)}
    \end{minipage}
    \begin{minipage}[t]{0.2\linewidth}\centering
        \begin{lstlisting}
            struct CustomerSlice<'a> {
                key: &'a i32,
                age: &'a i32,
                num_purchase: &'a i32,
                total_purchase: &'a f64,
                duration_spent: &'a f64, 
                duration_since: &'a f64,
                zip_code: &'a str,
                address: &'a str,
                country: &'a str, 
                state: &'a str,
                first_name: &'a str,
                last_name: &'a str,
                province: &'a str,
                comment: &'a str,
                order: &'a OrderSlice<'a>
            }
        \end{lstlisting}
      \medskip
      \centerline{(c)}
    \end{minipage}\hfill
    \begin{minipage}[t]{0.6\linewidth}\centering
        \begin{lstlisting}
            struct CustomerRc {
                key: Rc<i32>,
                age: Rc<i32>,
                num_purchase: Rc<i32>,
                total_purchase: Rc<f64>,
                duration_spent: Rc<f64>,
                duration_since: Rc<f64>,
                zip_code: Rc<String>,
                address: Rc<String>,
                country: Rc<String>,
                state: Rc<String>,
                first_name: Rc<String>,
                last_name: Rc<String>,
                province: Rc<String>,
                comment: Rc<String>,
                order: Rc<OrderRc>
            }
        \end{lstlisting}
      \medskip
      \centerline{(c)}
    \end{minipage}\hfill
    \caption{Representation of Customer objects Whose fields are different variable type: (a) CustomerOwned struct whose fields are all owned 
    (b) CustomerBorrowed struct whose fields are borrowed with reference (c) CustomerSlice struct whose fields are borrowed with slice for sequence value, otherwise reference}
    \label{fig:customer}
 \end{figure}

 \begin{figure}[htb]
    \begin{minipage}[t]{0.2\linewidth}\centering
        \begin{lstlisting}
            struct OrderOwned {
                order_id: i32,
                num_items: i32, 
                payment: f64,
                order_time: f64,
                title: String,
                comment: String
            } 
        \end{lstlisting}
      \medskip
      \centerline{(a)}
    \end{minipage}\hfill
    \begin{minipage}[t]{0.6\linewidth}\centering
        \begin{lstlisting}
            struct OrderBorrowed<'a> {
                order_id: &'a i32,
                num_items: &'a i32, 
                payment: &'a f64,
                order_time: &'a f64,
                title: &'a String,
                comment: &'a String
            }
        \end{lstlisting}
      \medskip
      \centerline{(b)}
    \end{minipage}
    \begin{minipage}[t]{0.2\linewidth}\centering
        \begin{lstlisting}
            struct OrderSlice<'a> {
                order_id: &'a i32,
                num_items: &'a i32, 
                payment: &'a f64,
                order_time: &'a f64,
                title: &'a str,
                comment: &'a str
            }
        \end{lstlisting}
      \medskip
      \centerline{(c)}
    \end{minipage}\hfill
    \begin{minipage}[t]{0.6\linewidth}\centering
        \begin{lstlisting}
            struct OrderRc {
                order_id: Rc<i32>,
                num_items: Rc<i32>,
                payment: Rc<f64>,
                order_time: Rc<f64>,
                title: Rc<String>,
                comment: Rc<String>
            }
        \end{lstlisting}
      \medskip
      \centerline{(c)}
    \end{minipage}\hfill
    \caption{Representation of Order objects Whose fields are different variable type: (a) OrderOwned struct whose fields are all owned 
    (b) OrderBorrowed struct whose fields are borrowed with reference (c) OrderSlice struct whose fields are borrowed with slice for sequence value, otherwise reference}
    \label{fig:order}
 \end{figure}

 \subsection{Wikipedia Data Sets}
 \label{sec:history}
 Wikipedia page data sets are used to perform document classification with KNN. 
 \(10^5\) pages are used for training data set, and 18724 pages are used for target.

 \subsection{Experimental Details}
 Our experiments are run on three types of VM instances on Google Cloud Platform: 
 n1-standard-1 which has 1 vCPU, 3.75 GB RAM, and 10 GB Standard persistent disk, 
 n1-standard-4 which has 2 vCPU, 15 GB RAM, and 10 GB Standard persistent disk, 
 n1-standard-4 which has 2 vCPU, 15 GB RAM, and 10 GB Standard persistent disk,  
 All of the experiments described are performed 5 times. 
 In this thesis, we present the average of 5 separate runs for each experiment.
\clearpage

### 2020-04-21 thesis_template/3_Body/model/access_own_reference_slice.tex
documented different variable and drop reference

\subsection{Concept}
This experiment is conducted to understand two questions. One is how different pointer types have impact to runtime performance.
The other is how initialization of Vec size has impact to runtime performance. 
In this experiment, we focus on owner, reference, and additionally slice as a pointer to sequence values. 
The memory representations of these pointer types are shown in figure.

The owner has a pointer pointing to the memory address of sequence values, length of the values, and capacity allocated to store additional values. 
The reference is a pointer that points to the owner. The slice is a pointer that points to memory address of sequence values and has value such as length of 
sequence values stored in the memory. Since they have different memory representation, an assumption is that it takes different time to access to the contents of the memory among these pointer types.

To evaluate this assumption, we use the three type of complex object.

@@

\begin{figure}[htb]
    \includegraphics[width=15cm]{rust_access_init_vs_noint.eps}
    \caption{Runtime of Access to Fields of Complex Object with Initialization vs without Initialization}
    \label{fig:rustaccessnoinit}
\end{figure}

The representation of these objects are shown in Figure~\ref{fig:customer} and Figure~\ref{fig:order}. 
\begin{figure}[htb]
    \begin{minipage}[t]{0.2\linewidth}\centering
        \begin{lstlisting}
            struct CustomerOwned {
                key: i32,
                age: i32,
                num_purchase: i32,
                total_purchase: f64,
                duration_spent: f64, 
                duration_since: f64,
                zip_code: String,
                address: String,
                country: String,
                state: String,
                first_name: String,
                last_name: String,
                province: String,
                comment: String, 
                order: OrderOwned
            }
        \end{lstlisting}
      \medskip
      \centerline{(a)}
    \end{minipage}\hfill
    \begin{minipage}[t]{0.6\linewidth}\centering
        \begin{lstlisting}
            struct CustomerBorrowed<'a> {
                key: &'a i32,
                age: &'a i32,
                num_purchase: &'a i32,
                total_purchase: &'a f64,
                duration_spent: &'a f64, 
                duration_since: &'a f64,
                zip_code: &'a String,
                address: &'a String,
                country: &'a String,
                state: &'a String,
                first_name: &'a String,
                last_name: &'a String,
                province: &'a String,
                comment: &'a String, 
                order: &'a OrderBorrowed<'a>
            }
        \end{lstlisting}
      \medskip
      \centerline{(b)}
    \end{minipage}
    \begin{minipage}[t]{0.2\linewidth}\centering
        \begin{lstlisting}
            struct CustomerSlice<'a> {
                key: &'a i32,
                age: &'a i32,
                num_purchase: &'a i32,
                total_purchase: &'a f64,
                duration_spent: &'a f64, 
                duration_since: &'a f64,
                zip_code: &'a str,
                address: &'a str,
                country: &'a str, 
                state: &'a str,
                first_name: &'a str,
                last_name: &'a str,
                province: &'a str,
                comment: &'a str,
                order: &'a OrderSlice<'a>
            }
        \end{lstlisting}
      \medskip
      \centerline{(c)}
    \end{minipage}\hfill
    \caption{Representation of Customer objects Whose fields are different variable type: (a) CustomerOwned struct whose fields are all owned 
    (b) CustomerBorrowed struct whose fields are borrowed with reference (c) CustomerSlice struct whose fields are borrowed with slice for sequence value, otherwise reference}
    \label{fig:customer}
 \end{figure}

 \begin{figure}[htb]
    \begin{minipage}[t]{0.2\linewidth}\centering
        \begin{lstlisting}
            struct OrderOwned {
                order_id: i32,
                num_items: i32, 
                payment: f64,
                order_time: f64,
                title: String,
                comment: String
            } 
        \end{lstlisting}
      \medskip
      \centerline{(a)}
    \end{minipage}\hfill
    \begin{minipage}[t]{0.6\linewidth}\centering
        \begin{lstlisting}
            struct OrderBorrowed<'a> {
                order_id: &'a i32,
                num_items: &'a i32, 
                payment: &'a f64,
                order_time: &'a f64,
                title: &'a String,
                comment: &'a String
            }
        \end{lstlisting}
      \medskip
      \centerline{(b)}
    \end{minipage}
    \begin{minipage}[t]{0.2\linewidth}\centering
        \begin{lstlisting}
            struct OrderSlice<'a> {
                order_id: &'a i32,
                num_items: &'a i32, 
                payment: &'a f64,
                order_time: &'a f64,
                title: &'a str,
                comment: &'a str
            }
        \end{lstlisting}
      \medskip
      \centerline{(c)}
    \end{minipage}\hfill
    \caption{Representation of Order objects Whose fields are different variable type: (a) OrderOwned struct whose fields are all owned 
    (b) OrderBorrowed struct whose fields are borrowed with reference (c) OrderSlice struct whose fields are borrowed with slice for sequence value, otherwise reference}
    \label{fig:order}
 \end{figure}

>>>>

This experiment is conducted to understand two questions. One is how different variable types have impact to runtime performance.
The other is how initialization of Vec size has impact to runtime performance. 
In this experiment, we focus on owner, reference, and slice as a variable of sequence values. 
Since these variables have different memory representation, there might be differences among time for access to actual values of each variables.

To evaluate this assumption, we use the three types of complex object: CustomerOwned, CustomerBorrowed and CustomerSlice. 
At first, we generate source Vecs for all fields, Vecs which contain all elements used for corresponding fields of objects.
For example, all of i32 elements used for key field in 1 million Customer object are stored in Vec$<$i32$>$ with 1 million i32 elements. 
Later, these i32 elements are moved to be owned or borrowed by by the object's fields.

Next, 1.5, 1.8, 2.5, 2.8, 3.5, and 3.8 million of Customer objects are created and stored in Vec. 
When a Customer Vec is created, whether size of Vec is initialized is controlled. 
Finally, serialization of Customer object is performed as an operation forcing program to access all of fields in the object.
This serialization is performed each Customer objects in the Vec. We measure total runtime to serialize all of Customer objects 
stored in Vec.

\subsection{Result}
\label{sec:history}
The result is shown in Figure~\ref{fig:rustaccessinit} and Figure~\ref{fig:rustaccessnoinit}.
Figure~\ref{fig:rustaccessinit} is comparison of the runtime performance among different Customer object types with Vec size initialization.
Figure~\ref{fig:rustaccessnoinit} is comparison in the same set of experiment except Vec size is not initialized. 
The blue, yellow, and green chats represent runtime of access to fields of CustomerOwned, CustomerBorrowed, and CustomerSlice objects respectively.

No matter Vec size initialized or not, differences of runtime for accessing objects among different object types are not remarkable. 
The percent increase in runtime of accessing to fields of from 1.5 to 3.8 million CustomerOwned objects is 121\% if the Vec is initialized, 554\% if the Vec is not initialized.

Considering the differences of percent increase of runtime whether the Vec is initialized or not, 
Figure\ref{fig:init_vs_noinit} the comparison of runtime access to fields of CustomerOwned object with and without Vec Initialization.
For 3.8 million of objects, the difference of runtime to access to the objects is 38745 secounds; program without Vec initialization is 60\% slower than one with the initialization.

@@

\begin{figure}[htb]
    \includegraphics[width=15cm]{rust_access_init_vs_noint.eps}
    \caption{Runtime of Access to Fields of Complex Object with Initialization vs without Initialization}
    \label{fig:init_vs_noinit}
\end{figure}

\subsection{Discussion}
\label{sec:history}
Difference of variable types does not have huge impact to runtime of accessing to actual value.
Even thought owner, reference , and slice have different memory representations, the access time to its value is 
close to each other. As shown in Figure~\ref{fig:own_ref_slice}, the representations of owner and slice are almost identical except slice does not have capacity for values.
Reference is pointer pointing to owner, so it has an additional step to access actual value. 
However, the result shows this additional step does not have huge impact for runtime to access memory region of the value.

Whether initializing Vec size results in disparity of runtime performance to access objects' fields. 
This is because when Vec uninitialized, the elements of Vec are allocated across different virtual memory pages.

### 2020-04-21 thesis_template/3_Body/model/drop_ref_rc.tex
documented different variable and drop reference

\subsection{Concept}
The goal of this experiment is to assess efficiency of different reference strategies in Rust, borrowing and reference counting.

By levaraging reference counting, a value can be shared like what borrowing plays the role in Rust programming. 
The difference is that reference counting checks number of reference pointing to the actual data and makes sure the data is not deleted 
until all the references are dereferenced. Using reference counting is sometimes preferable approach for developers, 
we do not have care about lifetime which is usually troublesome when we use borrowing approach many times in our code. 
However, the possible problem regarding to reference counting is the cost for tracking the number of references. 
Having this assumption, this experiment will show difference of behavior among reference countinig and borrowing.

In this experiment, CustomerBorrowed and CustomerRc in figure are used to see difference of dropping time among borrowing and reference counting. 
In the CustomerRc and OrderRc struct, all fields take reference counting (Rc$<$T$>$). Similaly to the experiment in the last section, 
vectors of CustomerRc and CustomerBorrowed are created and droped the elements one by one.

The result shows significant difference of dropping time among the two objects; deletion of CustomerRc is much slower than CustomerBorrowed. 
This is because reference counters which are feilds of CustomerRc have to check the number of reference pointing to the actual content and decide 
deallocate the memory or not. However, memory management and lifetime strategy of borrowing is already determined at compile time.

In this experiment, an assessment is conducted to valify whether there is difference between behavior of borrowing and reference counting. 
These two methods can be used in similar situation. Especially, when we want reference to an original variable and keep the original variable until all of references are dereferenced, 
we can have reference with both storategies, borrowing and reference counting. One advantage of using reference counting is that we do not have to pay attention to lifetime.
However, the assumption here is that use of reference counting might be computationally expenesive than borrowing, because reference counting has to track the number of reference pointing to the value. 
Considered this assumption, we assess difference of time for dropping reference using borrowing and reference counting strategies.

Similally to the last section, we implement struct whose fields are borrowing reference and reference counting and measure the time to drop the struct.

\subsection{Result}
The result shows the time to drop borrowing reference is significatly faster than reference counting. This says that we should use borrwoing storategies whenever high performance computation is critical.
\begin{figure}[htb!]
    \includegraphics[width=15cm]{rust_droptime_borring_rc.eps}
    \caption{Runtime for Droping to fields of Customer Object}
    \label{fig:Sampling}
\end{figure}

\subsection{Discussion}

>>>>

% In Rust programming, a reference is a pointer to reduce unnecessary movement of ownership. 
% One use case is where we pass reference to values to arguments of function. 
% If we pass owner to function, the owner is moved and the corresponding value is deallocated when the call of function ends. 
% To avoid this, the owner should always be returned. This may be an encumbrance while complex software system development. 
% By using reference, we do not have to worry about returning it, because reference is an additional pointer to an owned value. 
% Reference can be used for operation in the same way of owner, but also be moved without deallocation of value by keeping its owner live.

% Reference is useful to avoid movement of ownership. However, one needs to track its lifetime and explicitly includes it in code, 
% because Rust compiler cannot infer it. This can be another encumbrance. We can instead acquire multiple owners to single value by using Reference Counting (Rc). 
% By leveraging Rc, a value can be shared like what borrowing plays the role in Rust programming. 
% The difference is that Rc checks number of owner pointing to the actual data and makes sure the data is not deleted 
% until all the owners are dereferenced. Using Rc is sometimes preferable approach for developers especially when lifetime planning is extremely difficult.
% However, the possible problem regarding to Rc is the cost for tracking the number of references. 
% Having this assumption, this experiment will show difference of behavior among Rc and simple reference.

In this experiment, CustomerBorrowed and CustomerRc are used to see difference of dropping time among reference and Rc. 
In the CustomerRc and OrderRc struct, all fields take Rc (Rc$<$T$>$). Similarly to the experiment in the last section, 
sets of integer, float, and String vector are created and their elements are borrowed or reference counted to create CustomerBorrowed or CustomerRc objects.
The dropping of objects deletes references or Rcs used for fields of the objects. However, it does not deallocate values to which they are pointing. 
Therefore, the evaluated runtime of dropping objects only consists of dropping time of reference or Rc, but deallocation time.
We generated 1, 1.5, 2, 2.5, 3, 3.5 million of CustomerBorrowed and CustomerRc objects and performed drop one by one.

\subsection{Result}
Figure\ref{fig:rc_ref} shows comparison of runtime dropping CustomerBorrowed and CustomerRc objects. 
The result shows significant difference of dropping time among the two objects; deletion of CustomerRc is much slower than CustomerBorrowed. 
The runtime of dropping CustomerBorrowed is about 60 times faster than dropping CustomerRc.

\begin{figure}[htb!]
    \includegraphics[width=15cm]{rust_droptime_borring_rc.eps}
    \caption{Runtime for dropping Customer Object}
    \label{fig:rc_ref}
\end{figure}

\subsection{Discussion}
In this experiment, an assessment is conducted to verify whether there is difference between behavior of reference and Rc.
The reason why dropping Rc is much slower than dropping reference it that Rc requires runtime overhead to check some states of the variable, 
but when to drop reference is already determined at compile time. When dropping Rc, Rc has to check the number of variable pointing to the actual content and decide 
deallocate the memory or not. However, memory management and lifetime strategy of reference is already determined at compile time.
This determination of memory management strategy at compile increases runtime performance of dropping complex object constructed with reference type variable.
This may say that we should use reference whenever high performance computation is critical.

However, dealing with reference is sometimes cumbersome. Tracking lifetime of reference can be done easily in simple situation. 
But, if we have complex objects constructed with fields of reference, the lifetime tracking become extremely difficult. 
For example, constructing nested objects with reference fields require a developer to plan memory management with many lifetime symbols. 
Using reference counting eliminates developer's responsibility to specify lifetime of variables. 
This may ease and speed up development process, and increase understandability of codes.

Even though we have stack allocated values, such as i32 and f64, in Rc in our experiment, 
one should avoid wrapping stack allocated values in Rc. Wrapping value in Rc allocates heap memory so that allocating Rc$<$i32$>$ or Rc$<$f64$>$ unnecessarily use space of 
heap. Additionally, stack allocated values are usually easy to be copied. Therefore, developer does not have to even use reference; one can just copy the value. 
Copying value in Rust is to copy the original value and to assign the copy to new owner variable.
